---
title: "GHFC Co-op Reopening Checkin Desk Analysis"
author: "Josh Hadro"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(tsibble)

#old = read_csv('./data/processed_logs.csv')

new = read_csv('/Users/joshhadro/github/coop_reports/logs/logs/processed_logs.csv')

rescued_data = read_csv('/Users/joshhadro/github/coop_reports/logs/logs/rescued_july_2020_data.csv')

rescued_data_01_2021 = read_csv('/Users/joshhadro/github/coop_reports/logs/jan_2021_data/january_2021_rescued_data.csv')


df = full_join(new, rescued_data)
df = full_join(df, rescued_data_01_2021)

rounded <- df %>%
  mutate(round = round_date(df$checkin_dt, "5 seconds")) %>%
  group_by(round, checkin_mode, checkin_type) %>%
  ungroup() %>%
  distinct(round, checkin_mode, checkin_type, .keep_all = TRUE) %>%
  select(-round)


all = rounded %>%
  filter(checkin_dt > '2020-06-05', # Exclude sketchy first 1.5 days of data collection while bugs were being worked out
         #checkin_dt < today() # Exclude an incomplete day
         ) %>%
  mutate(weekday = wday(checkin_dt, label = TRUE, week_start = getOption("lubridate.week.start", 1)))


```

# Last January Weekend report

Here are the recorded numbers for the last few days:

Checkins Friday: `  36`    
Checkins Saturday: `68`    
Checkins Sunday:`   95`   


```{r out.width = '95%', echo=FALSE, message=FALSE }

all %>%
  filter(checkin_dt > '2021-01-29') %>%
  group_by(date = date(checkin_dt), checkin_type) %>%
  #group_by(date = date(checkin_dt)) %>%
  summarize(count = n()) %>%
  drop_na() %>%
  ggplot(aes(date, count, fill = checkin_type)) +
  #ggplot(aes(date, count)) +
  geom_bar(stat='identity', position = 'dodge') +
  scale_fill_brewer(palette="Set1") +
  theme(legend.position = 'bottom') + 
  labs(x = 'Date', y = "Number of checkins", fill='Checkin type') +
  #facet_wrap(~checkin_type)
  geom_text(aes( x = date,
                 y = count,
    label=count
    ),
    check_overlap = TRUE,
    stat= "identity",
    vjust = -.3,
           #size = 5,
           position = position_dodge(width = 0.9)
            ) 
```

### Biggest shopping days

Sunday, January 31 was the Co-ops largest recorded day in terms of checkins; as Amanda LP put it, "mostly because of snowstorm as well as PSFC being closed."

The other top 10 days are show below.

Interestingly, four of the top 10 checkin days have been in the last few weeks.

```{r out.width = '95%', echo=FALSE, message=FALSE }

library(kableExtra)


options(knitr.table.format = "html") 

all %>%
  group_by(Date = date(checkin_dt)) %>%
  summarize(`Total Checkins` = n()) %>%
  drop_na() %>%
  mutate(
    `Day of the week` = wday(Date, label = TRUE, abbr = FALSE),
    Note = case_when(
    #lead(GROWTH > 1000, default = F), 0.3, SIZE)
    Date == '2021-01-31' ~ "Sunday before major snowstorm",
    Date == '2020-11-22' ~ "Sunday before Thanksgiving",
    Date == '2020-09-13' ~ "Sunday after Labor Day",
    Date == '2021-01-30' ~ "Saturday before major snowstorm"
    )
      ) %>%
  arrange(-`Total Checkins`) %>%
  head(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```



# Recurring report

In June 2020 the Co-op reinstated its check-in procedure. As part of the IT group's work to revise the check-in application, we're able to capture a lot more usable data for analysis of who is coming into the store and when.

A server restart issue caused a logging error from June 26 - July 1, which is apparent in the figures below where there is missing data. 

(See the previous report [("GHFC Co-op Checkin Desk Analysis")](https://hadro.github.io/ghfc/ghfc_psfc_report) for a view into the data we were able analyze from mid-March 2019 through mid-March 2020, when the previous iteration of the check-in desk was in operation.)

N.B.: this data reflects only what was captured -- put another way, if people did not check in at the check-in desk or mis-entered their data, then those statistics would not, by definition, be captured here. That said, I have no reason to believe the figures here are under-representative in any significant way. Moreover, the improved UX of the new checkin desk makes it easy to check in, and I'm confident in the few days of data we have. 

## Shoppers per day

The following figure shows the number of shoppers per day starting June 05. Each dot represents the total number of check-ins on a single day. 

```{r out.width = '95%', echo=FALSE, message=FALSE }

# Time series by day

library(zoo)


all %>%
  group_by(date = date(checkin_dt)) %>%
  summarize(count = n(), 
            weekday = first(weekday)) %>% 
  ggplot(aes(date, count)) +
  geom_point()+geom_smooth()+
  scale_fill_brewer(palette="Set1") +
  theme(legend.position = 'bottom') + 
  labs(x = 'Date', y = "Number of shoppers", color='Shopper type')


```

### Shoppers per week, by type

Meanwhile, the revised check-in system allows for more granular faceting. For example, we now capture more specifically what type of visitor each check-in is associated with.

There are four possible values: 

- GHFC (Greene Hill Food Co-op member)
- PSFC (Park Slope Food Co-op member)
- Trial member
- Visitor


### Proportions of shopper types

```{r out.width = '95%', echo=FALSE, message=FALSE }

all %>%
  group_by(checkin_type) %>%
  mutate(checkin_type_ct = n()) %>%
  ungroup() %>%
  mutate(pct = (checkin_type_ct / n())) %>%
  ggplot(aes(checkin_type, fill = checkin_type))+
  geom_bar() +
  scale_fill_brewer(palette="Set1") +
  theme(legend.position = 'bottom') + 
  labs(title = "Proportion of shoppers",
       x = 'Checkin type', y = "Number of shoppers", fill='Checkin type') +
  geom_text(aes( x = checkin_type,
                 y = checkin_type_ct, 
    label=scales::percent(pct, accuracy = 1)
    ),
    check_overlap = TRUE,
    stat= "identity",
    vjust = -.3,
           #size = 5,
           position = position_dodge(width = 0.9)
            ) 
```

<!-- ### Totals by shopper type by week -->

<!-- ```{r out.width = '95%', echo=FALSE, message=FALSE } -->



<!-- all %>% -->
<!--   group_by(week = yearweek(checkin_dt), year = year(checkin_dt), checkin_type) %>% -->
<!--   #filter(n_distinct(weekday) >= 5) %>% -->
<!--   #select(-checkin_mode, -can_shop, -membership_status, -membership_start_date) %>% -->
<!--   mutate(week_type_ct = n(), -->
<!--                  week_start = floor_date(checkin_dt, unit = "week", week_start = getOption("lubridate.week.start", 3)), -->
<!--                  week_label = format(week_start, format=paste('%b %e %y')) -->
<!--            ) %>% -->
<!--   ungroup() %>% -->
<!--   group_by(week_label) %>% -->
<!--   mutate(pct = (week_type_ct / n())) %>% -->
<!--   ggplot(aes(x = reorder(week_label, week), group = checkin_type, fill = checkin_type)) + -->
<!--   geom_bar(aes(y = week_type_ct), position="dodge", stat = "identity")+ -->
<!--   #scale_y_continuous(labels=scales::percent)+ -->
<!--   scale_fill_brewer(palette="Set1") + -->
<!--   theme(legend.position = 'bottom') + -->
<!--   labs(x = 'Week of the Year', y = "Number of shoppers", fill='Checkin type') + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1))+ -->
<!--   facet_wrap(~checkin_type) -->

<!-- ``` -->


#### GHFC shoppers per day, by checkin type

To give you a sense of relative daily numbers, the figure below shows the total shoppers per day broken out by each of these four check-in types.

```{r out.width = '95%', echo=FALSE, message=FALSE }

# Time series by day

all %>%
  group_by(date = date(checkin_dt), checkin_type) %>%
  summarize(count = n()) %>%
  drop_na() %>%
  ggplot(aes(date, count, fill = checkin_type)) +
  geom_bar(stat='identity', position = 'dodge') +
  #geom_area(position = 'stack') +
  #geom_smooth() +
  scale_fill_brewer(palette="Set1") +
  theme(legend.position = 'bottom') + 
  labs(x = 'Date', y = "Number of shoppers", fill='Checkin type') +
  facet_wrap(~checkin_type)




```

#### GHFC shoppers per day, by membership plan

Likewise, now with the new check-in desk we can also look at how the various membership plans within the GHFC check-in type are represented over the same period.
In particular, how many folks are taking advantage of the Summer Special Zucchini Plan.

```{r out.width = '95%', echo=FALSE, message=FALSE }
all %>%
  group_by(date = date(checkin_dt), membership_type) %>%
  summarize(count = n()) %>%
  drop_na() %>%
  ggplot(aes(date, count, fill = membership_type)) +
  geom_bar(stat='identity', position = 'dodge') +
  #geom_area(position = 'stack') +
  #geom_smooth() +
  #scale_fill_brewer(palette="Set2") +
  theme(legend.position = 'bottom') +
  labs(x = 'Date', y = "Number of shoppers", fill='Membership type') +
  facet_wrap(~fct_rev(membership_type), nrow = 2)

```




### Average shoppers by day of the week and hour of the day

Finally, we can break down this data even further to look at hours within each given day. 6pm (18 on the figure below) is consistently the most popular shopping time on weekdays, while mid- to late-afternoon is consitently the most popular shopping time on weekends.

```{r out.width = '95%', echo=FALSE, message=FALSE, warning=FALSE }
#by hour

all %>%
  mutate(date = date(checkin_dt),
    hour = hour(checkin_dt),
         weekday = wday(checkin_dt, label = TRUE, week_start = 1)) %>%
  filter(hour > 9.5, hour < 22) %>%
  group_by(hour, date) %>%
  mutate(kind_hour_count = n()) %>%
  ungroup() %>%
  group_by(weekday, hour) %>%
  mutate(hour_avg = mean(kind_hour_count)) %>%
  ggplot(aes(hour, hour_avg, fill = 'green'))+
  geom_bar(stat="identity", position = "dodge") +
  facet_wrap(~weekday, ncol=7)+
  coord_equal()+
  scale_x_continuous(limits=c(9,21)) + scale_y_continuous(limits=c(0,12.5)) +
    scale_fill_brewer(palette="Set2") +
  theme(legend.position = 'none') +
  labs(x = 'Hour of the day', y = "Average number of shoppers", fill='Visitor type')

```


Or we can look at the same data by day and time, an also add in checkin type:

```{r out.width = '95%', echo=FALSE, message=FALSE, warning=FALSE }
#by hour

all %>%
  mutate(date = date(checkin_dt),
    hour = hour(checkin_dt),
         weekday = wday(checkin_dt, label = TRUE, week_start = 1)) %>%
  filter(hour > 9.5, hour < 22) %>%
  group_by(hour, date, checkin_type) %>%
  mutate(kind_hour_count = n()) %>%
  ungroup() %>%
  group_by(checkin_type, weekday, hour) %>%
  summarize(hour_avg = mean(kind_hour_count)) %>%
  ggplot(aes(hour, hour_avg, fill = checkin_type))+
  geom_bar(stat="identity", position = "dodge") +
  facet_grid(checkin_type~weekday, scales = 'free')+
  scale_x_continuous(limits=c(9,21)) + scale_y_continuous(limits=c(0,12.5)) +
    scale_fill_brewer(palette="Set1") +
  theme(legend.position = 'bottom') +
  labs(x = 'Hour of the day', y = "Average number of shoppers", fill='Visitor type')

```

That's it for now -- but as always, please let me know if there's a question you'd like to investigate based on this data or any other data, and I'll be happy to take a look.


<!-- ```{r out.width = '95%', echo=FALSE, message=FALSE, warning=FALSE } -->
<!-- #Calendar heatmap -->

<!-- require(devtools) -->

<!-- install_github("jbryer/makeR") -->
<!-- require(makeR) -->

<!-- library("calendarHeat") -->

<!-- calendarHeat(amznStock$date, amznStock$adjusted, ncolors = 99, color = "r2g", varname="AMZN Adjusted Close") -->

<!-- all %>% -->
<!--   mutate(date = date(checkin_dt), -->
<!--     hour = hour(checkin_dt), -->
<!--          weekday = wday(checkin_dt, label = TRUE, week_start = 1)) %>% -->
<!--   group_by(date) %>% -->
<!--   mutate(count = n()) -> temp  -->

<!-- calendarHeat(temp$date, temp$count, ncolors = 99, color = "r2b", varname="AMZN Adjusted Close") -->


<!-- ``` -->

<!-- ```{r out.width = '100%', echo=FALSE, message=FALSE, warning=FALSE } -->
<!-- #Calendar heatmap take 2 -->

<!-- by_day <- -->
<!--   all %>% -->
<!--   mutate( -->
<!--     #date_time = paste(`DTSTART-DATE`, `START-TIME`), -->
<!--     #date_time = mdy_hm(checkin_dt), -->
<!--     date_time = checkin_dt, -->
<!--     # get the date of the sunday prior to each appointment -->
<!--     created_day = floor_date(date_time, "day"), -->
<!--     wday = wday(date_time) -->
<!--   ) %>% -->
<!--   select(created_day, wday) %>% -->
<!--   na.omit() %>% # (I had a blank line in the csv file) -->
<!--   mutate( -->
<!--     wday = factor(wday, levels = 1:7, labels = c( -->
<!--       "Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat" -->
<!--     )), -->
<!--     # assign weeks and months for counting and labeling purposes -->
<!--     week = floor_date(created_day, "week"), -->
<!--     week = as_date(week), -->
<!--     month = floor_date(created_day, "month"), -->
<!--     month = as_date(month) -->
<!--   ) %>% -->
<!--   group_by(week, wday, month) %>% -->
<!--   # month is unnecessary to group by  -->
<!--   # but it's a good trick to keep that variable! -->
<!--   count() %>% -->
<!--    mutate( -->
<!--     n = as.factor(as.numeric(cut_number(freq,5))), -->
<!--     #n = factor(as.character(n), levels = c("1", "2", "3", "4")), -->
<!--     # add a 0 level to the 1:4 counts and move it to the front -->
<!--     #n = fct_expand(n, "0"), -->
<!--     #n = fct_relevel(n, "0") -->
<!--   ) -->

<!-- # match the format of the axis labels -->
<!-- months <- seq(min(by_day$month), max(by_day$month), by = "month") -->
<!-- month_labels <- strftime(months, "%b") -->
<!-- # create the text annotations -->
<!-- less <- grid::textGrob("Less", gp = grid::gpar(fontsize = 10, col = "#767676")) -->
<!-- more <- grid::textGrob("More", gp = grid::gpar(fontsize = 10, col = "#767676")) -->
<!-- p <- ggplot(by_day) + -->
<!--   aes(week, fct_rev(wday), fill = n) + -->
<!--   geom_tile(width = 7, height = 1) + -->
<!--   # decided to make the "spaces" lines  -->
<!--   # instead of actual spaces between grey tiles -->
<!--   geom_hline(yintercept = seq(.5, 7.5, 1), col = "white", size = .85) + -->
<!--   geom_vline( -->
<!--     xintercept = seq( -->
<!--       as_date("2018-01-01"), as_date("2018-12-31"), by = "week" -->
<!--     ) + 2.5, -->
<!--     col = "white", size = .85 -->
<!--   ) + -->
<!--   # the expand = F argument tells it to use those exact limits, no extra -->
<!--   coord_fixed( -->
<!--     ratio = 7,  -->
<!--     xlim = c(min(by_day$month) + 2.5, max(by_day$month) + 25),  -->
<!--     expand = FALSE -->
<!--     ) + -->
<!--   labs(x = NULL, y = NULL) + -->
<!--   scale_x_date( -->
<!--     expand = c(0, 0), breaks = months, labels = month_labels, position = "top" -->
<!--     ) + -->
<!--   scale_y_discrete(labels = c("", "Fri", "", "Wed", "", "Mon", "")) + -->
<!--   scale_fill_gradient(low="red", high="green") + -->
<!--   scale_fill_manual( -->
<!--     limits = levels(by_day$n), -->
<!--     values = c("#EBEDF0", "#C6E48B", "#7BC96F", "#239A3B", "#196127"), -->
<!--     name = NULL -->
<!--    ) + -->
<!--   theme( -->
<!--     # ugh so much trial and error to get these numbers: -->
<!--     legend.position = c(.904, -.405), -->
<!--     legend.justification = "bottom", -->
<!--     panel.grid = element_blank(), -->
<!--     panel.background = element_rect(fill = "#EBEDF0"), -->
<!--     axis.ticks.y = element_blank(), -->
<!--     axis.ticks.x = element_blank(), -->
<!--     axis.text.x = element_text(hjust = 0.5, color = "#767676", size = 10), -->
<!--     axis.text.y = element_text(color = "#767676", size = 10), -->
<!--     plot.margin = margin(4, 5, 4, 4), -->
<!--     legend.key.size = unit(10, "pt"), -->
<!--     legend.text = element_blank(), -->
<!--     legend.spacing.x = unit(.05, "cm"), -->
<!--     plot.title = element_text(hjust = 0, vjust = 0) -->
<!--   ) + -->
<!--   #ggtitle(paste0(nrow(cal) - 1, " appointments in the last year")) + -->
<!--   guides(fill = guide_legend(nrow = 1)) + -->
<!--   annotation_custom(less, -->
<!--     xmin = as_date("2018-10-25"), -->
<!--     xmax = as_date("2018-10-25"), ymin = -2, ymax = -1 -->
<!--   ) + -->
<!--   annotation_custom(more, -->
<!--     xmin = as_date("2018-12-20"), -->
<!--     xmax = as_date("2018-12-20"), ymin = -2, ymax = -1 -->
<!--   ) -->
<!-- # this is necessary to get the annotations outside the plotting area to print -->
<!-- gt <- ggplot_gtable(ggplot_build(p)) -->
<!-- gt$layout$clip[gt$layout$name == "panel"] <- "off" -->
<!-- grid::grid.draw(gt) -->



<!-- ``` -->


<!-- ```{r out.width = '100%', echo=FALSE, message=FALSE, warning=FALSE } -->


<!-- library(plyr) -->
<!-- library(plotly) -->

<!-- amznStock = by_day -->
<!-- amznStock$date = amznStock$created_day -->

<!-- amznStock$weekday = as.POSIXlt(amznStock$date)$wday #finding the day no. of the week -->
<!-- amznStock$weekday = wday(amznStock$date, week_start = 1 ) #finding the day no. of the week -->

<!-- amznStock$weekdayf<-factor(amznStock$weekday,levels=rev(1:7),labels=rev(c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")),ordered=TRUE) # converting the day no. to factor -->

<!-- amznStock$monthf<-factor(month(amznStock$date),levels=as.character(1:12),labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"),ordered=TRUE) # finding the month -->

<!-- amznStock$yearmonth<- factor(as.yearmon(amznStock$date)) # finding the year and the month from the date. Eg: Nov 2018 -->

<!-- amznStock$week <- as.numeric(format(amznStock$date,"%W")) # finding the week of the year for each date -->

<!-- amznStock<-ddply(amznStock,.(yearmonth),transform,monthweek=1+week-min(week)) # normalizing the week to start at 1 for every month -->


<!-- ggplot(amznStock, aes(monthweek, weekdayf, fill = amznStock$freq)) +  -->
<!--     geom_tile(colour = "white") + facet_grid(year(amznStock$date)~monthf) +  -->
<!--   scale_fill_gradient2(low="purple", mid = "white", high="green", midpoint = mean(amznStock$freq)) +   -->
<!--   #  scale_colour_gradient2(midpoint = 25) + -->
<!--   xlab("Week of Month") + ylab("") + ggtitle("Time-Series Calendar Heatmap: Checkins per day") + labs(fill = "Checkins")+ -->
<!--   coord_fixed(    ratio =1,  -->
<!-- ) -->

<!-- ``` -->